---
title: "Vignette of reading the fars_read function"
author: "JC"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette of reading the fars_read function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Vignette description

## The 'fars_read' function
The 'fars_read' function returns a table of the class 'data frame' when the input is a filename from the Fatality Analysis Reporting System (FARS) of the US National Highway Traffic Safety.

### Details 
When the file name exists, the fars_read function will be applied havng the following steps.
First, the file will be evaluated whether it is a csv file with comma as separator by the 'suppressMessages' function from the base package.
If the file is a csv, the file will be read in using the 'read_csv' function from the readr package and assigned as 'data'.
Then, this 'data' will be converted in a table having a 'data frame' class, using the 'tbl_df' function from the dplyr package.



```{r reading_farsFile}
fars_read <- function(filename) {
  if(!file.exists(filename))
    stop("file '", filename, "' does not exist")
  data <- suppressMessages({
    readr::read_csv(filename, progress = FALSE)
  })
  dplyr::tbl_df(data)
}

```


## The 'make_filename' function
The 'make_filename' function creates a filename, having a charactor vector containing a formatted combination
of text and the 'year' variable.

### Details
When the year is given, the year is converted into a integer by 'as.integer' from the base package, and
assigned as 'year'. Then, the 'year' variable will be used in the C function 'sprintf' from the base package.

```{r creating_farsFilename}
make_filename <- function(year) {
  year <- as.integer(year)
  createdFilename <- sprintf("accident_%d.csv.bz2", year)
  systemFilename <- system.file('extdata', createdFilename, package = 'farsReadTest')
  systemFilename
}

```
