---
title: "Vignette of reading the fars_read function"
author: "JC"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette of reading the fars_read function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Vignette description

## The 'fars_read' function
The 'fars_read' function returns a table of the class 'data frame' when the input is a filename from the Fatality Analysis Reporting System (FARS) of the US National Highway Traffic Safety.

### Details 
When the file name exists, the fars_read function will be applied havng the following steps.
First, the file will be evaluated whether it is a csv file with comma as separator by the 'suppressMessages' function from the base package.
If the file is a csv, the file will be read in using the 'read_csv' function from the readr package and assigned as 'data'.
Then, this 'data' will be converted in a table having a 'data frame' class, using the 'tbl_df' function from the dplyr package.



```{r reading_farsFile}
fars_read <- function(filename) {
  if(!file.exists(filename))
    stop("file '", filename, "' does not exist")
  data <- suppressMessages({
    readr::read_csv(filename, progress = FALSE)
  })
  dplyr::tbl_df(data)
}

```


## The 'make_filename' function
The 'make_filename' function creates a filename, having a charactor vector containing a formatted combination
of text and the 'year' variable.

### Details
When the year is given, the year is converted into a integer by 'as.integer' from the base package, and
assigned as 'year'. Then, the 'year' variable will be used in the C function 'sprintf' from the base package.

```{r creating_farsFilename}
make_filename <- function(year) {
  year <- as.integer(year)
  createdFilename <- sprintf("accident_%d.csv.bz2", year)
  systemFilename <- system.file('extdata', createdFilename, package = 'farsReadTest')
  systemFilename
}

```


## The 'fars_read_years' function
The 'fars_read_years' function returns a list of each element of the 'year' variable which appears in the read-in file.

### Details
Within the 'fars_read_years' function, a 'lapply' function from the base package is used. The parameters for
the 'lapply' function are 'years' and the 'make_filename' function (defined above).
Within the 'lapply' function, the 'make_filename' function will be first applied on the 'year' variable and assigned
this as 'file'. Then, the 'tryCatch' function is called and evaluates its content, containing an expression and the function(e)
which implies an error. The expression is composed of:
    1) the 'fars_read' function of a file as 'dat'
    2) data manipulation of the 'dat' file: adding a new variable 'year' using the 'mutate' function of the dplyr package,
                                             followed by selecting the variables 'MONTH' and 'year' from the 'dat' file
When the year is valid, a list of valid years and months will be returned.

```{r reading_multipleYears}
fars_read_years <- function(years) {
  lapply(years, function(year) {
    file <- make_filename(year)
    tryCatch({
      dat <- fars_read(file)
      dplyr::mutate(dat, year = YEAR) %>%
        dplyr::select(MONTH, year)
    }, error = function(e) {
      warning("invalid year: ", year)
      return(NULL)
    })
  })
}

```


## The 'fars_summarize_years' function
The 'fars_summarize_years' function returns a table with a summary of the 'years' parameter.

### Details
After applying the 'fars_read_years' function on the parameter 'years', a 'dat_list' object is made.
Lists of different years will be bound by rows using the 'bind_rows' function of the dplyr package, followed by:
   1) the 'group_by' function of the dplyr package, aggregating by the 'year' and 'MONTH' variables
   2) after aggregation, the number of rows 'n' is counted using the 'summarize' function from the dplyr package
   3) the different elements in the 'year' column is then spread over multiple columns using the 'spread'
      function from the dplyr package, together with the number of rows 'n'

```{r summarising_years}
fars_summarize_years <- function(years) {
  dat_list <- fars_read_years(years)
  dplyr::bind_rows(dat_list) %>%
    dplyr::group_by(year, MONTH) %>%
    dplyr::summarize(n = n()) %>%
    tidyr::spread(year, n)
}

```


## The fars_map_state' function
The 'fars_map_state' function returns a plot with points which represent the occurring coordinates in the given state
number and year in a 'accident_%.csv' file.

### Details
Within the 'fars_map_state' function, the followings handling are executed:
   1) a filename will be created using the 'make_filename' function (defined previously) by given the 'year' parameter
   2) the created filename is then used as input for the 'fars_read' function in order to return a table of the class
     'data.frame'. This table is named as 'data'
   3) the 'state.num' parameter is converted to an integer (by 'as.integer')
   4) a search for the 'state.num' is performed in a the column 'STATE' in the 'data' table which is made unique. When the
      state number s valid, the 'filter' function of the dplyr package will be applied to obtain all the information of that
      state number in the 'data' table. This is assigned as 'data.sub' (=subset of the 'data' table). Within the 'data.sub',
      the followings are executed:
        a) converting the values of LONGITUD variable greater than 900 into 'NA'
        b) converting the values of LATITUDE variable greater than 90 into 'NA'
        c) finally, the 'data.sub' data will be used to create a graph using:
             - the 'map' function of the maps package for drawing lines and polygons using the 'state' database
               with the range of the 'LATITUDE' variable as the Y-axis and the range of the 'LONGITUD' variable as the 'X-axis'
               to create the raster/plot. Both having the NA values removed when executing this step.
            - the 'points' function of the graphics package to add the coordinates (x = LONGITUD, y = LATITUDE) as points
               (using pch = 46) in the raster/plot.

```{r mapping}
fars_map_state <- function(state.num, year) {
  filename <- make_filename(year)
  data <- fars_read(filename)
  state.num <- as.integer(state.num)

  if(!(state.num %in% unique(data$STATE)))
    stop("invalid STATE number: ", state.num)
  data.sub <- dplyr::filter(data, STATE == state.num)
  if(nrow(data.sub) == 0L) {
    message("no accidents to plot")
    return(invisible(NULL))
  }
  is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
  is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
  with(data.sub, {
    maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
              xlim = range(LONGITUD, na.rm = TRUE))
    graphics::points(LONGITUD, LATITUDE, pch = 46)
  })
}

```
